---
title: "projec2_new_file"
author: "Rose Eluvathingal Muttikkal"
date: "2023-07-13"
output: pdf_document
---

```{r}
library(ggplot2)
library(faraway)
library(zipcodeR)
library(tidyverse)
library(gridExtra) 
```

Data wrangling
```{r}
Data <- read.csv("kc_house_data.csv")
Data <- Data<-Data[, -c(1:2)] 
```

Collapse categorical variables to manageable levels

bed
```{r}
Data <- Data %>% 
  mutate(high_bed =ifelse(bedrooms>4, "yes", "no")) 
```

bathrooms
```{r}
Data <- Data %>% 
  mutate(high_bathrms =ifelse(bathrooms>4, "yes", "no")) 
```

floors
```{r}
Data <- Data %>% 
  mutate(high_floor =ifelse(floors>2, "yes", "no")) 
```

view
```{r}
Data <- Data %>% 
  mutate(good_view =ifelse(view>2, "yes", "no")) 
```

condition 
```{r}
Data <- Data %>% 
  mutate(good_condition =ifelse(condition>2, "yes", "no")) 
```

yr_renovated (0 not, other yes)
```{r}
Data <- Data %>% 
  mutate(renovated =ifelse(yr_renovated==0, "no", "yes")) 
```

zipcode to cities
```{r}
#for(i in 1:length(Data$zipcode)){
  #Data$city[i] <- reverse_zipcode(Data$zipcode[i])$major_city
#}
#Data$city <- unlist(Data$city)
```


north, south, east, west
```{r}
Data <- Data %>% 
  mutate(is_east =ifelse(long>-121.786309, "yes", "no")) %>% 
  mutate(is_north =ifelse(lat>47.498010, "yes", "no")) 
```
Factorize categorical variables
```{r}
Data$high_bed = factor(Data$high_bed)
Data$high_bathrms = factor(Data$high_bathrms)
Data$high_floor = factor(Data$high_floor)
Data$waterfront = factor(Data$waterfront)
Data$good_view = factor(Data$good_view)
Data$good_condition = factor(Data$good_condition)
Data$renovated = factor(Data$renovated)
Data$isNorth = factor(Data$is_north)
Data$isEast = factor(Data$is_east)
```

```{r}
ystar<-log(Data$price) 
Data<-data.frame(Data,ystar) 
```
Split into train + test
```{r}
set.seed(6021) 
##evenly split data into train and test sets 
sample.data<-sample.int(nrow(Data), floor(.50*nrow(Data)), replace = F) 
train<-Data[sample.data, ] 
test<-Data[-sample.data, ] 
```

Quant data set
```{r}
D1<-train%>% 
  dplyr::select(ystar) 
D2<-train%>% 
  dplyr::select(yr_built) 
D3<-train%>% 
  dplyr::select(grade) 
D4<-train%>% 
  dplyr::select(sqft_living) 
D5<-train%>% 
  dplyr::select(bedrooms) 
D6<-train%>% 
  dplyr::select(sqft_basement) 
D7<-train%>% 
  dplyr::select(sqft_lot15) 
D8<-train%>% 
  dplyr::select(sqft_lot) 

quant<-bind_cols(D1, D2, D3, D4, D5, D6, D7, D8)
head(quant) 
  
```

QUANT --> Multicolinearity
```{r}
round(cor(quant[,-1]),3)
```
QUANT --> VIFS 
```{r}
all<-lm(ystar~., data=quant) 
faraway::vif(all) 
```

Whole data set
```{r}
D5<-train%>% 
  dplyr::select(renovated) 
D6<-train%>% 
  dplyr::select(good_condition) 
D7<-train%>% 
  dplyr::select(good_view) 
D8<-train%>% 
  dplyr::select(waterfront) 
D9<-train%>% 
  dplyr::select(high_bathrms) 
D10<-train%>% 
  dplyr::select(high_bed)
D11<-train%>% 
  dplyr::select(is_north)
D12<-train%>% 
  dplyr::select(is_east)

whole<-data.frame(quant, D5, D6, D7, D9, D10, D11, D12)
head(whole) 
```

WHOLE - mine
Stepwise --> OUR INTIAL FULL MODEL
```{r}
##intercept only model 
regnull <- lm(ystar~1, data=whole) 
##model with all predictors 
regfull <- lm(ystar~., data=whole) 
step(regnull, scope=list(lower=regnull, upper=regfull), direction="both")
```
using model selection criteria - not useful
```{r}
allreg <- leaps::regsubsets(price ~., data=whole, nbest=1) 
summary(allreg) 
which.min(summary(allreg)$bic) 
```
```{r}
coef(allreg, which.min(summary(allreg)$bic)) 
```

WHOLE - WHOLE - not using
Step
```{r}
##intercept only model 
regnull <- lm(price~1, data=train) 
##model with all predictors 
regfull <- lm(price~., data=train) 
step(regnull, scope=list(lower=regnull, upper=regfull), direction="both")
```
QAUNT ONLY
Step
```{r}
##intercept only model 
regnull <- lm(price~1, data=quant) 
##model with all predictors 
regfull <- lm(price~., data=quant) 
step(regnull, scope=list(lower=regnull, upper=regfull), direction="both")
```
CAT ONLY
CAT data set
```{r}
D3<-train%>% 
  select(city) 
D4<-train%>% 
  select(renovated) 
D5<-train%>% 
  select(good_condition) 
D6<-train%>% 
  select(good_view) 
D7<-train%>% 
  select(waterfront) 
D8<-train%>% 
  select(high_floor) 
D9<-train%>% 
  select(high_bathrms) 
D10 <-train%>% 
  select(price) 

cat<-bind_cols(D10, D8,D3, D4, D5, D6, D7, D9)
head(cat) 
```
Step - not using (CAT)
```{r}
##intercept only model 
regnull <- lm(price~1, data=cat) 
##model with all predictors 
regfull <- lm(price~., data=cat) 
step(regnull, scope=list(lower=regnull, upper=regfull), direction="both")
```
Checking AICs of all candidate models - may not be usefl
```{r}
model1 <- lm(price ~ sqft_living + city + grade + yr_built + good_view + high_bathrms + lat + bedrooms + long + sqft_basement + high_bed + renovated, data = train)

model2 <- lm(price ~ sqft_living + lat + yr_built + grade + bedrooms + long, data = train)

model3 <- lm(price ~ lat + yr_built + grade + sqft_living + city + good_view + high_bathrms, data = train)

model4 <- lm(price ~ sqft_living + sqft_lot15 + grade + high_bathrms + bedrooms + city, data = train)

model5 <- lm(price ~ sqft_living + sqft_lot15 + grade + high_bathrms + high_bed + city, data = train)
```
No long + lat - may not be useful
```{r}
model1 <- lm(price ~ sqft_living + city + grade + yr_built + good_view + high_bathrms + bedrooms + sqft_basement + renovated, data = train) #high bed repeating

model2 <- lm(price ~ sqft_living + yr_built + grade + bedrooms, data = train)

model3 <- lm(price ~ yr_built + grade + sqft_living + city + good_view + high_bathrms, data = train)

model4 <- lm(price ~ sqft_living + sqft_lot15 + grade + high_bathrms + bedrooms + city, data = train)

model5 <- lm(price ~ sqft_living + sqft_lot15 + grade + high_bathrms + high_bed + city, data = train)
```

```{r}
library(AICcmodavg)

#define list of models
models <- list(model1, model2, model3, model4, model5)

#specify model names
mod.names <- c('1', '2', '3', '4', '5')

#calculate AIC of each model
aictab(cand.set = models, modnames = mod.names)
```
```{r}
model1 <- lm(price ~ sqft_living + city + grade + yr_built + good_view + high_bathrms + bedrooms + sqft_basement + renovated, data = train)
summary(model1) 
```
```{r}
##check dummy coding 
contrasts(Data$city) 
```
Residual plots --> to help determine log transformation
```{r}
result<-lm(price~sqft_living + is_north + good_view + grade + yr_built + high_bathrms + bedrooms + renovated + high_bed + sqft_lot15 + is_east + sqft_basement + sqft_lot, data=train) 
par(mfrow = c(2, 2)) 
plot(result) 
```
```{r}
library(MASS) ##to use boxcox function 
MASS::boxcox(result, lambda = seq(-0.2, 0.2, 1/10))
```
```{r}
##transform y and then regress ystar on x 
result.ystar<-lm(ystar~sqft_living + is_north + good_view + grade + yr_built + high_bathrms + bedrooms + renovated + high_bed + sqft_lot15 + is_east + sqft_basement + sqft_lot, data=train) 
par(mfrow = c(2, 2)) 
plot(result.ystar) 
```

Assessing full model
```{r}
result<-lm(ystar ~ grade + is_north + sqft_living + yr_built + 
    good_view + sqft_lot + good_condition + renovated + is_east + 
    high_bathrms + high_bed + sqft_basement, data=train) 
summary(result)
```

Checking for interactions - not needed
GRADE
```{r}
options(scipen=999)
ggplot2::ggplot(train, aes(x=grade, y=ystar, color=is_north))+ 
  geom_point()+ 
  geom_smooth(method=lm, se=FALSE)+ 
  labs(title="Scatterplot of Log Transformed Price against Grade, by Direction (North/South)")+ 
  geom_text(label=rownames(train)) 
```
```{r}
ggplot2::ggplot(train, aes(x=grade, y=ystar, color=is_east))+ 
  geom_point()+ 
  geom_smooth(method=lm, se=FALSE)+ 
  labs(title="Scatterplot of Log Transformed Price against Grade, by Direction (East/West)")+ 
  geom_text(label=rownames(train)) 
```
```{r}
ggplot2::ggplot(train, aes(x=grade, y=ystar, color=good_view))+ 
  geom_point()+ 
  geom_smooth(method=lm, se=FALSE)+ 
  labs(title="Scatterplot of Log Transformed Price against Grade, by View")+ 
  geom_text(label=rownames(train)) 
```
```{r}
ggplot2::ggplot(train, aes(x=grade, y=ystar, color=good_condition))+ 
  geom_point()+ 
  geom_smooth(method=lm, se=FALSE)+ 
  labs(title="Scatterplot of Log Transformed Price against Grade, by Condition")+ 
  geom_text(label=rownames(train)) 
```
```{r}
ggplot2::ggplot(train, aes(x=grade, y=ystar, color=renovated))+ 
  geom_point()+ 
  geom_smooth(method=lm, se=FALSE)+ 
  labs(title="Scatterplot of Log Transformed Price against Grade, by Renovation Status")+ 
  geom_text(label=rownames(train)) 
```

```{r}
##model with interaction 
result.int<-lm(ystar ~ grade + is_north + sqft_living + yr_built + 
    good_view + sqft_lot + good_condition + renovated + is_east + 
    high_bathrms + high_bed + sqft_basement + grade*is_north + grade*good_view + grade*good_condition + grade*renovated + grade*is_east + grade*high_bathrms + grade*high_bed + sqft_living*is_north + sqft_living*good_view + sqft_living*good_condition + sqft_living*renovated + sqft_living*is_east + sqft_living*high_bathrms + sqft_living*high_bed + yr_built*is_north + yr_built*good_view + yr_built*good_condition + yr_built*renovated + yr_built*is_east + yr_built*high_bathrms + yr_built*high_bed + sqft_basement*is_north +  sqft_basement*good_view + sqft_basement*good_condition + sqft_basement*renovated + sqft_basement*is_east + sqft_basement*high_bathrms + sqft_basement*high_bed, data = train) 
summary(result.int) 
```
General Linear F-test 

Fail to reject null, Use reduced
```{r}
reduced<-lm(ystar ~ grade + is_north + sqft_living + yr_built + 
    good_view + sqft_lot + good_condition + renovated + is_east + 
    high_bathrms + high_bed + sqft_basement + grade*is_north + grade*good_condition + grade*high_bed + sqft_living*is_north + sqft_living*good_view + sqft_living*good_condition + sqft_living*high_bathrms + sqft_living*high_bed + yr_built*is_north + yr_built*good_view + yr_built*good_condition + yr_built*renovated + yr_built*is_east + sqft_basement*is_north +  sqft_basement*good_view + sqft_basement*high_bathrms + sqft_basement*high_bed, data = train) 
##general linear F test to compare reduced model with full model 
anova(reduced, result.int) 
```
NEW MODEL (considering interactions) - not important
```{r}
summary(reduced)
```
General linear F-test --> MLR
```{r}
reduced2<-lm(ystar ~is_north + sqft_living + yr_built + 
    good_view + sqft_lot + good_condition + renovated + is_east + high_bed + sqft_basement + grade*is_north + grade*good_condition + sqft_living*is_north + sqft_living*good_view + sqft_living*good_condition + yr_built*is_north + yr_built*good_view + yr_built*good_condition + yr_built*renovated + yr_built*is_east + sqft_basement *is_north +  sqft_basement*good_view + sqft_basement*renovated, data = train) 
anova(reduced, reduced2) 
```

leverage - 
```{r}
hii<-lm.influence(result)$hat ##leverages 
ext.student<-rstudent(result) ##ext studentized res
n<-nrow(train) 
p<- 13 
leverage <- sort(hii[hii>2*p/n], decreasing = TRUE)
leverage_10 <- leverage[1:10]
leverage_10 
```
externally studentized - outliers
```{r}
outliers <- sort(ext.student[abs(ext.student)>3], decreasing = TRUE)
outliers_10 <- outliers[1:10]
outliers_10
```
DFFITS
```{r}
DFFITS<-dffits(result) 
DFFITS <- sort(DFFITS[abs(DFFITS)>2*sqrt(p/n)], decreasing = TRUE)
DFFITS_10 <- DFFITS[1:10]
DFFITS_10
```

```{r}
##compute yhat and yhat(i) for the influential obs anove 
y<-train$ystar 
yhat<-y-result$res 
del.res<-result$res/(1-hii) 
##deleted residual 
yhat.i<-y-del.res 
##yhat(i) ##compare yhat, yhat(i), and compute their difference for the states found above
c_bind <- cbind(yhat,yhat.i, yhat-yhat.i)[abs(DFFITS)>2*sqrt(p/n),] 
c_bind <- as.data.frame(c_bind)
c_bind<-c_bind%>% 
  arrange(desc(V3)) 
c_bind
```
Mean Squared Error - M1
```{r}
data <- data.frame(pred = predict(result, newdata = test), actual = test$ystar)
mean((data$actual - data$pred)^2)
```
```{r}
summary(result)
```

Residual plot
```{r}
par(mfrow = c(2, 2)) 
plot(result) 
```





