---
title: "proj2"
author: "Rose Eluvathingal Muttikkal"
date: "2023-07-11"
output: pdf_document
---
```{r}
library(ggplot2)
library(faraway)
library(zipcodeR)
library(tidyverse)
```


Data wrangling
```{r}
Data <- read.csv("kc_house_data.csv")
Data <- Data<-Data[, -c(1:2)] 

set.seed(6021)
sample.data <- sample.int(nrow(Data), floor(0.50*nrow(Data)), replace=F)
train <- Data[sample.data,]
test <- Data[-sample.data,]
```

SLRs --> identify numeric relationship - living, above, living15, lot15 (looks like real linear relationship)
```{r}
options(scipen=999)
dp4 <- ggplot2::ggplot(train, aes(x=yr_built,y=price))+ 
  geom_point()+ 
  geom_smooth(method = "lm", se=FALSE)+ 
  labs(x="Year Built", y="Price", title="Scatterplot of Price against Year Built")
dp4
```
Check multicollinearity 

bathrooms --> sqft_living
sqft_living --> sqft_above, sqft_liviing15
sqft_lot --> sqft_lot15
```{r}
quant <- train[, -c(6:10)] 
quant <- quant[, -c(9:12)] 
#quant <- quant[, -c(7:8)] 
#quant <- quant[, -c(7:8)] 
round(cor(quant[,-1]),3) 
```
VIFs (only quant) --> too highly correlated to compute, combine multiple into one, get rid of neighbors

```{r}
all<-lm(price~., data=quant) 
faraway::vif(all) 
```
Redo VIF + correlation matrix after getting rid of neighbors info
Redo VIF + correlation matrix after combining
Redo VIF + correlation matrix after combining + getting rid of neighbord

Model selection - All (model 1)

COLLAPSING --> categorical predictors
bedrooms 
```{r}
ggplot2::ggplot(train, aes(x=bedrooms,y=price))+ 
  geom_point()+ 
  geom_smooth(method = "lm", se=FALSE)+ 
  labs(x="Bedrooms", y="Price", title="Scatterplot of Price against Bedrooms")
ggplot2::ggplot(train, aes(x=bathrooms,y=price))+ 
  geom_point()+ 
  geom_smooth(method = "lm", se=FALSE)+ 
  labs(x="Bathrooms", y="Price", title="Scatterplot of Price against Bathrooms")
ggplot2::ggplot(train, aes(x=floors,y=price))+ 
  geom_point()+ 
  geom_smooth(method = "lm", se=FALSE)+ 
  labs(x="Floors", y="Price", title="Scatterplot of Price against Floors")
```

bathrooms
floors
view
condition
```{r}
train <- train %>% 
  mutate(good_condition =ifelse(condition>2, "yes", "no")) 
```

yr_renovated (0 not, other yes)
```{r}
#data <- train %>% 
  #filter(yr_renovated!=0) 
  
#ggplot2::ggplot(data, aes(x=yr_renovated,y=price))+ 
  #geom_point()+ 
  #geom_smooth(method = "lm", se=FALSE)+ 
  #labs(x="Year Renovated", y="Price", title="Scatterplot of Price against Year Renovated")

train <- train %>% 
  mutate(renovated =ifelse(yr_renovated==0, "no", "yes")) 
```

zipcode
```{r}
for(i in 1:length(train$zipcode)){
  train$city[i] <- reverse_zipcode(train$zipcode[i])$major_city
}

train$city <- unlist(train$city)
```

```{r}
write.csv(train, "/Users/rosee.m./Desktop/bookmarks/linear_models_ds/lm_project2/city.csv", row.names=TRUE)
```


--factored all variables/new variable categories

- get rid of zipcode, lat, long, yr_renovated, 
```{r}
##make discrete categories
train$bedrooms_cat = factor(train$bedrooms)
train$bathrooms = factor(train$bathrooms)
train$floors = factor(train$floors)

train$waterfront = factor(train$waterfront)
train$view = factor(train$view)
train$good_condition = factor(train$good_condition)
train$renovated = factor(train$renovated)
train$city = factor(train$city)
```

Forward
```{r}
##intercept only model 
regnull <- lm(price~1, data=train) 
##model with all predictors 
regfull <- lm(price~., data=train) 
step(regnull, scope=list(lower=regnull, upper=regfull), direction="forward") 
```

Backward
```{r}

```

Stepwise
```{r}

```


Model selection - Quant (model 2)

Forward
```{r}
##intercept only model 
regnull <- lm(price~1, data=quant) 
##model with all predictors 
regfull <- lm(price~., data=quant) 
step(regnull, scope=list(lower=regnull, upper=regfull), direction="forward") 
```

Backward
```{r}
##intercept only model 
regnull <- lm(price~1, data=quant) 
##model with all predictors 
regfull <- lm(price~., data=quant) 
step(regfull, scope=list(lower=regnull, upper=regfull), direction="backward")
```

Stepwise
```{r}
##intercept only model 
regnull <- lm(price~1, data=quant) 
##model with all predictors 
regfull <- lm(price~., data=quant) 
step(regnull, scope=list(lower=regnull, upper=regfull), direction="both")
```


Model selection - Quant with new variables (model 3)

Forward
Backward
Stepwise


Add categorical
Check for interactions
